\vspace{-15mm}\fontsize{16pt}{10pt}\selectfont\textbf{Facial Expression Recognition with\\ Convolutional Neural Networks}

### 2018-04-30 doc/paper.tex
phrasing

\section{Related work}
In the last years, great progress was made on the problem of automatically detecting facial expression from images. Some researchers have attempted to derive emotions from detecting individual muscle movements in images of faces by identifying muscles and their degree of activation \cite{lien98}. However, manually extracting these relevant features from the training data requires expert knowledge or complex algorithms. In Convolutional Neural Networks (CNNs), features can be learned from the training data implicitly and only the image pixels are required as input for predictions. With this convenience, CNNs have been used successfully for various image recognition tasks, including facial expression recognition \cite{dumas01}.
%------------------------------------------------

@@

% - CNN robustness to occlusion, distortions and geometric transformations.
The final model scored an accuracy of ~75\% on a test-set of 3557 samples. In live predictions with a webcam, the classifier has shown to be accurate. However, predictions for the classes $disgust$ and $fear$ are hard to reproduce. This is most likely because the seven classes of emotion in the dataset are highly imbalanced. For instance disgust only has 500 samples while happy has much more. This results in biased predictions with higher probabilities of the majority classes. In order to balance the training dataset, data augmentation could be used \cite{krizhevsky12}: To increase the number of training examples for a given class, the images could be flipped horizontally or rotated in multiple random angles. Given more time for this project, I would like to have experimented with such techniques of data augmentation and also with  more complex model architectures.

>>>>

\section{Related work}
In the last years, great progress was made on the problem of automatically detecting facial expression from images. Some researchers have attempted to derive emotions from detecting individual muscle movements in images of faces by identifying muscles and their degree of activation \cite{lien98}. However, manually extracting the muscle activations from a set of images requires expert knowledge or complex algorithms. With Convolutional Neural Networks (CNNs), training data only consists of images along with matching class labels that describe the images. Relevant features, such as a smile that implies happiness, are learned implicitly and do not require manual extraction. CNNs have been used successfully for various image recognition tasks, including facial expression recognition \cite{dumas01}.
%------------------------------------------------

@@

% - CNN robustness to occlusion, distortions and geometric transformations.
The final model scored an accuracy of ~75\% on a test-set of 3557 samples. In live predictions with a webcam, the classifier has shown to be accurate in most of the predictions. However, predictions for the classes $disgust$ and $fear$ are hard to reproduce. This is most likely because the seven classes of emotion in the dataset are highly imbalanced. For instance disgust only has about 500 samples while happy has ~7000. This results in biased predictions with higher probabilities of the majority classes. In order to balance the training dataset, data augmentation could be used \cite{krizhevsky12}: To increase the number of training examples for a given class, the images could be flipped horizontally or rotated in multiple random angles. Given more time for this project, I would like to have experimented with such techniques of data augmentation and also with  more complex model architectures.
